---
title: "Assignment 2"
author: "Insert My Name"
date: "2024-09-25"
output: html_document # you can change to other output format if you want
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(Stat2Data)
data("Hawks")
```

# 1. Data Wrangling

## 1.1 (Q1)   

<!-- Include your answers for each question under a second level heading eg ## 1.1 (Q1) -->
```{r include=TRUE}
hSF <- Hawks %>% filter(Species=="RT",Weight>=1000) %>% select("Wing","Weight","Tail")
print(head(hSF,10))
```

## 1.1 (Q2)
 **there are 3 variables in hSF, show it how many columns it has**
 
 **There are 4 examples nrow(), dim(), length(rownames()), summary() or str()**
 
 **obervations: print(nrow(hSF)) 398**
 
 **dim(hSF) 398 3 dim(hSF)[1] 398 dim(hSF)[2] 3**
 
 **length(rownames(hSF)) 398**
 
 **summary(hSF) overview of the data, str(hSF) all the data and the number of rows**

## 1.2 (Q1)
```{r include=TRUE}
sorthSF <- hSF %>% arrange(Wing)
print(sorthSF %>% head(10))
```


....

# 2. Random experiments, events and sample spaces, and the set theory

....

# 3. Probability theory

## 3 (Q1)

## 3 (Q2)

.....



