---
title: "Assignment 2"
author: "Guanjie Liang"
date: "2024-09-27"
output: html_document # you can change to other output format if you want
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(Stat2Data)
data("Hawks")
```

# 1. Data Wrangling

## 1.1 (Q1)   

<!-- Include your answers for each question under a second level heading eg ## 1.1 (Q1) -->
```{r include=TRUE}
hSF <- Hawks %>% filter(Species=="RT",Weight>=1000) %>% select("Wing","Weight","Tail")
print(head(hSF,10))
```


## 1.1 (Q2)
 **there are 3 variables in hSF, show it how many columns it has**
 
 **There are 4 examples nrow(), dim(), length(rownames()), summary() or str()**
 
 **obervations: print(nrow(hSF)) 398**
 
 **dim(hSF) 398 3 dim(hSF)[1] 398 dim(hSF)[2] 3**
 
 **length(rownames(hSF)) 398**
 
 **summary(hSF) overview of the data, str(hSF) all the data and the number of rows**


## 1.2 (Q1)
```{r include=TRUE}
sorthSF <- hSF %>% arrange(Wing)
print(sorthSF %>% head(10))
```


## 1.3 (Q1)
```{r include=TRUE}
species_code <- c("CH","RT","SS")
species_name_full <- c("Cooper's", "Red-tailed", "Sharp-shinned")
hawkSpeciesNameCodes <- data.frame(species_code, species_name_full)
print(hawkSpeciesNameCodes)
```


## 1.3 (Q2)
```{r include=TRUE}
hawksFullName <- hawkSpeciesNameCodes %>% rename(Species = species_code)
hawksFullName <- left_join(Hawks, hawksFullName) %>% select(-Species) %>% rename(Species = species_name_full)
print(hawksFullName %>% head(5))
```


## 1.3 (Q3)
```{r include=TRUE}
print(hawksFullName %>% select(Species, Wing, Weight) %>% head(7))
```


## 1.4 (Q1)
```{r include=TRUE}
hawksWithBMI <- Hawks %>% mutate(bird_BMI = 1000*Weight/Wing^2) %>% select(Species, bird_BMI) %>% arrange(desc(bird_BMI))
print(hawksWithBMI %>% head(8))
```


## 1.5 (Q1)
```{r include=TRUE}
hawksFullNameSum <- hawksFullName %>% group_by(Species) %>% summarize(num_rows=n(),
                                                                      mn_wing=mean(Wing, na.rm=TRUE),
                                                                      nd_wing=median(Wing,na.rm=TRUE),
                                                                      t_mn_wing=mean(Wing,min(0.1),na.rm=TRUE),
                                                                      b_wt_ratio=max(Wing/Tail, na.rm=TRUE))
print(hawksFullNameSum)
```


## 1.5 (Q2)
```{r include=TRUE}
hawksFullNameNA <- hawksFullName %>% group_by(Species) %>% summarize(Wing=sum(is.na(Wing)),
                                                                     Weight=sum(is.na(Weight)),
                                                                     Culmen=sum(is.na(Culmen)),
                                                                     Hallux=sum(is.na(Hallux)),
                                                                     Tail=sum(is.na(Tail)),
                                                                     StandardTail=sum(is.na(StandardTail)),
                                                                     Tarsus=sum(is.na(Tarsus)))
print(hawksFullNameNA)
```



# 2. Random experiments, events and sample spaces, and the set theory


## 2.1 (Q1)
**A Random experiment is a procedure (real or imagined) which:**
**1. has a well-defined set of possible outcomes**
**2. could (at least in principle) be repeated arbitrary many times**

**An event is a set of possible outcomes of an experiments**
**An event is any subset of the sample space, including the empty set and the sample space itself.**

**A sample space is the set of all possible outcomes of interest for a random experiment**


## 2.1 (Q2)
**event: {1,2}**

**sample space: {(1,1),(1,2),(1,3),(1,4),(1,5),(1,6),…,(6,6)}**

**total number of different events: $2^{6\times6}=2^{36}$** 

**Yes, the empty set is considered an event. It represents the impossible event**


## 2.2 (Q1)
**1. $A \cup B = \{1,2,3,4,6\} \quad A \cup B=\{1,2,3,4,5,6\}$ **

**2. $A \cap B = \{2\} \quad A \cap B =\{\}$**

**3. $A \backslash B=\{1,3\} \quad A \backslash B=\{1,2,3\}$**

**4. A and B are not disjoint, A and C are disjoint**

**5. Yes, B and $A \backslash B$  are disjoint**

**6. two sets: $\{\{1,2,3\},\{4,5,6\}\}$ <br>   three sets:$\{\{1,2\},\{3,4\},\{5,6\}\}**


## 2.2 (Q2)
**1. A**

**2. empty**

**3. $A^c=\Omega\backslash A, B^c=\Omega\backslash B \quad because \quad A \subseteq B, then \quad B^c \subseteq A^c$**

**4. $\cup_{k=1}^K A_{k}^c$**

**5. $(A \cup B)^c=\Omega \backslash (A \cup B)=(\Omega \backslash A) \cap (\Omega \backslash B)=A^c \cap B^c$**

**6. $\cap_{k=1}^K A_{k}^c$**


## 2.2 (Q3)
**$|E|=2^K$**


## 2.2 (Q4)
**1. empty set: $\emptyset$**

**2. $S_1 \cup S_2 \cup S_3 \cup S_4 = A_1 \cup A_2 \cup A_3 \cup A_4$**

**$\quad S_1 \cap S_2 = \emptyset \quad \quad S_2 \cap S_3 = \emptyset$**

**$\quad S_1 \cap S_3 = \emptyset \quad \quad S_2 \cap S_4 = \emptyset$**

**$\quad S_1 \cap S_4 = \emptyset \quad \quad S_3 \cap S_4 = \emptyset$**

**So $S_1,S_2,S_3,S_4 \quad form \quad a\quad partition \quad of \quad A_1 \cup A_2 \cup A_3 \cup A_4$**


## 2.2 (Q5)
**1. $1_{A^c}(w) = 1 - 1_A(w)$**

**2. $\Omega$**

**3. **

**Step 1: $1_{(A \cap B)^c}=1-1_{(A \cap B)}=1-1_A \cdot 1_B$**

**Step 2: $1_{A^c \cup B^c}=1_{A^c}+1_{B^c}-1_{A^c} \cdot 1_{B^c}=(1-1_A)+(1-1_B)-(1-1_A)(1-1_B)=1-1_A \cdot 1_B$**

**So $(A \cap B)^c = A^c \cup B^c$**


## 2.2 (Q6)
**the real number between 0 and 1 is infinite**


# 3. Probability theory

## 3.1 (Q1)
\[
P(x) = \begin{cases}
0, \quad A = \emptyset \\
0.5, \quad A = {a} \\
0.1, \quad A = {b} \\
0.4, \quad A = {c} \\
0.6, \quad A = {a,b} \\
0.9, \quad A = {a,c} \\
0.5, \quad A = {b,c} \\
1, \quad A = {a,b,c}
\end{cases}
\]


## 3.1 (Q2)
**1. $\mathbb{P}(A) \geq 0 \quad for \quad any \quad event \quad A$**

$\mathbb{P}(\emptyset)=0,\mathbb{P}({0})=1-q,\mathbb{P}({0,1})=1$

**2. $\mathbb{P}(\Omega)=\mathbb{P}({0,1})=1$**

**3. $\mathbb{P}(\cup_{i=1}^\infty A_i)=\mathbb{P}({0})+\mathbb{P}({1})=1$**

**$\sum_{i=1}^\infty \mathbb{P}(A_i)=\mathbb{P}({0})+\mathbb{P}({1})=1$**


## 3.2 (Q1)
**$\mathbb{P}(\cup_{i=1}^n A_i)=\mathbb{P}({A_1})+...+\mathbb{P}({A_n})$**

**$\sum_{i=1}^n \mathbb{P}(A_i)=\mathbb{P}({A_1})+...+\mathbb{P}({A_n})$**


## 3.2 (Q2)
**$\mathbb{P}(S) \cup \mathbb{P}(S^c) = \Omega, \quad \mathbb{P}(S) \cap \mathbb{P}(S^c) = \emptyset$**

**$So \quad \mathbb{P}(S^c)=1-\mathbb{P}(S)$**


## 3.2 (Q3)
**S_1={1,2},S_2={2,3},S_3={3,4}**


## 3.2 (Q4)
**Draw a diagram**


